### 2024-12-04 manuscript.tex
Update on Overleaf.

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution allows for parameterizations of
convection and clouds to be avoided, relying instead on explicitly resolved
processes.  Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station during
2010--2021 using a ground-based lidar simulator.  ICON and the reanalyses
underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON
and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially
explained by their lifting condensation levels being too high.  The reanalyses
strongly underestimate near-surface clouds or fog.  MERRA-2 tends to
underestimate cloud occurrence at all heights. Less stable conditions are the
most problematic for ICON and the reanalyses.  In daily cloud cover, ICON and
the reanalyses tend to be about 1 and 2 oktas clearer, respectively. Compared
to radiosondes, potential temperature is accurate in the reanalyses, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes were not enough
to address the model cloud biases.

@@

nextGEMS is a project \citep{nextgems} focused on the
research and development of GSRMs at multiple modelling centres and
universities in Europe.  The project also develops GSRM versions of the
Icosahedral Nonhydrostatic Weather and Climate Model (ICON), the Integrated
Forecasting System (IFS), and their ocean components at eddy-resolving
resolutions: ICON-O coupled with ICON and Finite-Element/volumE Sea ice-Ocean
Model (FESOM) and Nucleus for European Modelling of the Ocean (NEMO) coupled
with IFS.  The project has so far produced ICON and IFS simulations with three
development versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results match the observations better than the reanalyses in this metric.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution allows for parameterizations of
convection and clouds to be avoided, relying instead on explicitly resolved
processes, except for, among others, microphysics and turbulence.  Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station during
2010--2021 using a ground-based lidar simulator.  ICON and the reanalyses
underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON
and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially
explained by their lifting condensation levels being too high.  The reanalyses
strongly underestimate near-surface clouds or fog.  MERRA-2 tends to
underestimate cloud occurrence at all heights. Conditions of weak stability are associated with the largest errors in ICON and the reanalyses.  In daily cloud cover, ICON and
the reanalyses tend to be about 1 and 2 oktas clearer, respectively. Compared
to radiosondes, potential temperature is accurate in the reanalyses, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except microphysics) were not enough
to address the model cloud biases.

@@

nextGEMS is a project \citep{nextgems} focused on the
research and development of GSRMs at multiple modelling centres and
universities in Europe.  The project also develops GSRM versions of the
Icosahedral Nonhydrostatic Weather and Climate Model (ICON; \cite{hohenegger2023}), the Integrated
Forecasting System (IFS), and their ocean components at eddy-resolving
resolutions: ICON-O coupled with ICON and Finite-Element/volumE Sea ice-Ocean
Model (FESOM) and Nucleus for European Modelling of the Ocean (NEMO) coupled
with IFS.  The project has so far produced ICON and IFS simulations with three
development versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results ch the observations better than the reanalyses in this metric.
